\Huge \textcolor{titleTextColor}{ \textsf{ \fontsize{130}{100}\selectfont \textbf{GRASS}\,{\firalight GIS}: A General-purpose Geospatial Research Tool } }

### 2018-11-29 grassposter/2018_AGU_Geospatial_Research_Tool/poster.tex
poster: GRASS GIS: A General-purpose Geospatial Research Tool
git-svn-id: https://svn.osgeo.org/grass/grass-promo@73722 15284696-431f-4ddb-bdfa-cd5b030d7da7

% License: CC BY-SA
% Authors: See authors below and see also acknowledgement for authors of some images or research

\documentclass[25pt, margin=0mm, innermargin=15mm, blockverticalspace=15mm, colspace=15mm, subcolspace=8mm]{tikzposter}
\geometry{paperwidth=82in,paperheight=41in}

@@

@@

\definecolor{titleTextColor}{HTML}{009000}
\definecolorpalette{grassColorPalette} {
  \definecolor{colorOne}{HTML}{419041}
  % \definecolor{colorTwo}{HTML}{cccccc}
  \definecolor{colorTwo}{HTML}{dddddd}
  \definecolor{colorThree}{HTML}{F1B52D}
  % \definecolor{colorThree}{HTML}{EFA126}
}

\usetheme{Rays}
\usecolorstyle[colorPalette=grassColorPalette]{Britain}

\title{
\Huge
\textcolor{titleTextColor}{
\textsf{\textbf{
\fontsize{85}{60}\selectfont
How innovations thrive in GRASS GIS
}}
}
}

@@

\titlegraphic{
\begin{minipage}{0.3\linewidth}
\includegraphics[height=\grasslogoheight]{grass}
\hspace{2em}
\includegraphics[height=\grasslogoheight]{osgeo_project}
\end{minipage}
\hfill
\begin{minipage}{0.20\linewidth}
\setlength{\baselineskip}{120pt}
\begin{flushright}
\includegraphics[height=\instlogoheight]{ncstate}
~
\includegraphics[height=\instlogoheight]{ctu_prague}
~
\includegraphics[height=\instlogoheight]{dwd}
~
\includegraphics[height=\instlogoheight]{mundialis}
~
\includegraphics[height=\instlogoheight]{fem_cri}
~
\includegraphics[height=\instlogoheight]{iwmi}
~
\includegraphics[height=\instlogoheight]{ec_jrc}
~
\includegraphics[height=\instlogoheight]{eurac}
\end{flushright}
\end{minipage}
\vspace{-\grasslogoheight}
}

@@

\author{
V\'{a}clav Petr\'{a}\v{s}\inst{1},
Yann Chemin\inst{2},
Martin Landa\inst{3},
Thomas Leppelt\inst{4},
Pietro Zambelli\inst{5},
Luca Delucchi\inst{6},
Margherita Di Leo\inst{7},
S\"{o}ren Gebbert,
and
Markus Neteler\inst{8}
}
\institute{
\large
\instlist{1}North Carolina State University, USA (wenzeslaus@gmail.com, vpetras@ncsu.edu);
\instlist{2}IWMI, Sri Lanka;
\instlist{3}FCE CTU in Prague, Czech Republic;
\instlist{4}DWD, Germany;\\
\instlist{5}EURAC Research, Institute for Renewable Energy, Italy;
\instlist{6}Fondazione Edmund Mach, Research and Innovation Centre, Italy;
\instlist{7}European Commission, JRC, Italy;
\instlist{8}mundialis GmbH \& Co. KG, Germany;
}

\hypersetup
{
    pdfauthor={V. Petras, Y. Chemin, M. Landa, S. Gebbert, P. Zambelli, M. Neteler, M. Di Leo},
    pdfsubject={},
    pdftitle={How innovations thrive in GRASS GIS},
    pdfkeywords={GIS, algorithms, methods, preservation, science, reproducibility}
}

@@

\newcommand{\blocktitlewrap}[1]{\textsf{\textbf{\huge#1}}}
% it is not possible (?) to change block title in the class, using wrapper
% the command introduced using:
%   sed -i 's/\\block{\([^}]*\)}/\\block{\\blocktitlewrap{\1}}/g' main.tex

@@

\begin{document}
\maketitle[width=0.92\textwidth]
% \maketitle
% \addlogo[north west]{(2,-1)}{9cm}{images/Grass_GIS}
%Please insert your institution logo here
% \addlogo[north east]{(-2,-2.5)}{4cm}{images/logo_FEM_CRI}
% \addlogo[north east]{(-2,-5.5)}{4cm}{images/NC_State_Seal}
% \addlogo[north east]{(-8,-2.5)}{4cm}{images/Logo_cvut}
% \addlogo[north east]{(-8,-6.5)}{4cm}{images/IWMI_logo}
% \addlogo[north east]{(-2,-10.5)}{4cm}{images/logo_ec-jrc}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Highlights}}
{
% \setlength{\parskip}{0.3ex}

\renewcommand{\labelitemi}{\textcolor{gray}{$\bullet$}\hspace{0.5ex}}
\newcommand{\blocksectiontitle}[1]{\bigskip\textbf{\textcolor{gray}{\textsf{#1}}}}

\blocksectiontitle{Poster topic highlights}

\begin{itemize}
 \item Algorithms and models included in GRASS GIS remain available long term \citep{chemin2015grass}.
 \item Analytical tools are not limited to one domain but spread across many fields.
 \item New tools can be built based on functionality or code of the existing ones
       regardless of the particular domain of problems they belong to.
 \item Both the functionality and the code are evaluated
       by the community of users and developers in different fields and scales.
% continuous automated tests (Petras, 2014 \cite{Petras2014}),
\end{itemize}

\blocksectiontitle{General GRASS GIS highlights}

\begin{itemize}
%  \item Community supports users and new developments through online
%  \item long-term moving forward
 \item The GRASS GIS development team takes care of interface and operating system related changes
       in the code provided by scientists and contributors.
 \item Cost estimated by Black Duck Open Hub is over \$30,000,000.
 % https://www.openhub.net/p/grass_gis/estimated_cost
 \item The free, libre and open source license used in GRASS GIS
       not only allows users to reduce GIS software costs to minimum,
       but also enables organizations to deploy cloud-based solutions without any license restrictions.
 \item GRASS GIS is used both directly and as a backend in other projects
       such as QGIS and R.
\end{itemize}

% oholoh hours

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Lidar Data Processing}}{

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Acknowledgements}}{

\newcommand{\listhspace}{\hspace{0.005\linewidth}}
\newcommand{\listlogowidth}{0.10\linewidth}
\newcommand{\listtextwidth}{0.82\linewidth}

\begin{minipage}{\listlogowidth}
\centering
\includegraphics[width=0.5\linewidth]{grass}
\end{minipage}
\listhspace
\begin{minipage}{\listtextwidth}
We acknowledge all members of the GRASS GIS community both users and developers.
We would like to thank Markus Metz, Anna Petrasova, Stepan Turek, and Radim Blazek.
\end{minipage}

\begin{minipage}{\listlogowidth}
\includegraphics[width=\linewidth]{osgeo}
\end{minipage}
\listhspace
\begin{minipage}{\listtextwidth}
Open Source Geospatial Foundation (OSGeo)
supports the collaborative development of open source geospatial software.
GRASS GIS is a OSGeo project and
OSGeo provides infrastructure for project
websites, mailing lists and source code management.
\end{minipage}

\begin{minipage}{\listlogowidth}
\includegraphics[width=\linewidth]{google}
\end{minipage}
\listhspace
\begin{minipage}{\listtextwidth}
Initial development of \gmodule{i.segment} module as well as several other developments
were done as a part of the Google Summer of Code project.
Google provides financial support to students and organizations participating in the Google Summer of Code project.
\end{minipage}

\vspace{0.2cm}

\textcolor{gray}{
\hrulefill
}

\vspace{0.1cm}

\newcommand{\qrcodesize}{0.05\linewidth}

% qrencode http://grass.osgeo.org -o qr_grass.eps -t EPS
% epspdf -b qr_grass.eps qr_grass.pdf

\begin{center}
\begin{tabular}{c}

% \hspace{5mm}

\begin{minipage}{\qrcodesize}
\includegraphics[width=\textwidth]{./images/qr_grass.pdf}
\end{minipage}
~
\begin{minipage}{0.15\linewidth}
\small {\href{http://grass.osgeo.org}{\nolinkurl{grass.osgeo.org}}}
\end{minipage}

\begin{minipage}{0.1\linewidth}
\href{http://creativecommons.org/licenses/by-sa/4.0/}{\includegraphics[width=\textwidth]{ccbysa}}
\end{minipage}
~
\begin{minipage}{0.35\linewidth}
\small This poster is licensed under a Creative Commons Attribution-ShareAlike 4.0 International License.
\end{minipage}

\end{tabular}
\end{center}

\vspace{-0.08cm}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Water, Floods and Erosion}}{

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Image Segmentations}}{
\gamodule{r.smooth.seg} (formerly \gmodulenolink{r.seg})
was created by Alfonso Vitti \citep{vitti2008free, vitti2012mumford}
and implemented a piece-wise smooth approximation of the original image
according to \cite{mumford1989optimal} and \cite{march1997variational}
which can be used to reduce noise in the original image.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\column{0.25}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Topology, Cleaning, Overlays, Attributes}}{
Besides basic vector analysis tools such as \gmodule{v.buffer}
and \gmodule{v.overlay},
suite of functionality for topology checking and cleaning is available
through \gmodule{v.build} and \gmodule{v.clean} modules.
The vector cleaning tools are particularly advantageous considering the use of non-topological exchange formats.
The \gmodule{v.clean} module was introduced in 2002 and several different contributors
extended its functionality with last large set of improvements included in GRASS GIS 7.0.
For that release all topological cleaning tools have been optimized
with regard to processing speed, robustness, and system requirements.
The processing speed of the \gmodule{v.clean} module was substantially improved as well.

\begin{center}
\begin{minipage}{0.9\linewidth}
\begin{center}
\includegraphics[width=.3\textwidth]{topo_original}
~
\includegraphics[width=.3\textwidth]{topo_errors}
~
\includegraphics[width=.3\textwidth]{topo_fixed}
Original vector imported without cleaning (left),
identified errors (middle),
automatically topologically corrected vector (right)
\end{center}
\end{minipage}
\end{center}

% v.in.ascii input= output=imported format=standard
% v.build -e map=imported error=build_errors
% v.clean -c input=imported output=clean error=cleaning_errors tool=snap,rmdangle,rmbridge,chbridge,bpol,prune threshold=5
% d.vect map=build_errors color=255:33:36 fill_color=none width=5 icon=basic/point size=30
% d.vect map=cleaning_errors color=255:33:36 fill_color=none width=5 icon=basic/point size=30
% d.vect map=imported color=56:16:108 fill_color=205:64:113 width=5

An experimental module \gamodule{v.feature.algebra} (formerly \gmodulenolink{v.mapcalc}) from 2002
% actually, the original by Radim Blazek (Jun 11, 2002) seems to be more similar to the current v.mapcalc
% it was using Vect_overlay_str_to_operator Vect_overlay
% it was replaced by code by Christoph Simon in Aug 7, 2002
was replaced by Python for GRASS GIS C libraries (PyGRASS) in GRASS GIS 7.0.
However, new \gamodule{v.mapcalc} module is available with richer syntax and functionality.

Attribute processing and queries in GRASS GIS can take an advantage of latest developments
in database management systems (DBMSs) as several DBMSs are supported most notably SQLite and PostgreSQL.
A PostGIS database can be connected including its geometry and topology.
In parallel to that native support the OGC Simple Features is available.

The GRASS GIS 7 releases come with faster vector processing backend
and more efficient internal vector format;
both disk and memory requirements were reduced.
The new spatial index performs queries faster (more than 10 times for large vectors compared to GRASS GIS 6).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Spatio-Temporal Data Analysis}}{
The time dimension was introduced in GRASS GIS 7.0 for vector maps, rasters, and 3D rasters
which transformed GRASS GIS into a fully-featured temporal GIS \citep{Gebbert20141, gebbert2015grass}.
Time series of map layers are managed in space time datasets, a new data type in GRASS GIS,
and are still accessible also as individual map layers.
Based on the GRASS GIS Temporal Framework Python application programming interface (API),
more than 50 modules were implemented to manage, analyze, process
and visualize space time datasets.
More than 100,000 map layers can be now handled efficiently in GRASS GIS.
%
Example usages of this functionality include
analysis of the
European Climate Assessment \& Dataset ECA\&D \citep{Haylock2008_climate_series}
and temperate climate zone in the European Union identification \citep{Gebbert20141}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Vector Network Analysis}}{
In 2003 Radim Blazek introduced shortest path analysis (\gmodule{v.net.path}),
traveling salesman (\gmodule{v.net.salesman}),
and several other modules for vector network analysis (staring with \gmodule{v.net}).
Over the years, the number of available algorithms increased to over 15 modules
including, for example, centrality measures and connected components added in 2009 by Daniel Bundala.
In 2014 Stepan Turek implemented turns support into all relevant vector network modules.
Also since GRASS GIS 7.0 all vector network analysis tools provide fine control over node costs.
% combination with r.cost/r.walk workflows \citep{Petrasova2014}

\centering
\begin{minipage}{0.5\linewidth}
\begin{center}
\includegraphics[width=\textwidth]{network}
Shortest path between two points in the Wake County road network (NC, USA)
\end{center}
\end{minipage}

\block{\blocktitlewrap{Landscape Structure}}{
A set of modules for multiscale analysis of landscape structure was added in 1992
by \cite{baker1992r}, who developed the \gmodulenolink{r.le} model similar to
FRAGSTATS \citep{mcgarigal1995fragstats}.
The modules were gradually improved to become \gmodule{r.li} in 2006.
Further development continued, with a significant
increase in speed \citep{tracrli} and a new interactive user interface.
\cite{rocchini2013calculating} used \gmodule{r.li} modules to implement
high level tool for calculating landscape diversity.

\vspace*{0.5cm}

\begin{minipage}{0.5\linewidth}
\includegraphics[width=\textwidth, trim={0 180 0 0}, clip]{diversity_classes}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\textwidth, trim={0 180 0 0}, clip]{diversity_shannon}
\end{minipage}

\vspace*{1ex}

Landuse classes and derived landscape diversity according to Shannon index near Charlotte, NC, USA
% r.diversity input=development_2006 prefix=diversity alpha=0.5 size=65
% r.li.shannon input="development_2006" config="conf_diversity_65.0" output="diversity_shannon_size_65.0"
}

@@

@@

>>>>

% License: CC BY-SA
% Authors: See the authors below and see also acknowledgement for authors of some images or research

\documentclass[25pt, margin=0mm, innermargin=25mm, blockverticalspace=25mm, colspace=25mm, subcolspace=8mm]{tikzposter}
\geometry{paperwidth=63in,paperheight=42in}

@@

% Fira Sans, GRASS GIS branding sans serif font
\usepackage{FiraSans}
\renewcommand*\oldstylenums[1]{{\firaoldstyle #1}}

% EB Garamond, GRASS GIS branding serif font
% note that EB Garamond does not have bold
\usepackage[cmintegrals,cmbraces]{newtxmath}
\usepackage{ebgaramond-maths}

% might be needed for both font packages
\usepackage[T1]{fontenc}

% uncomment for all sans serif
% \renewcommand{\familydefault}{\sfdefault}

@@

\definecolor{titleTextColor}{HTML}{000000}
\definecolorpalette{grassColorPalette} {
  \definecolor{colorOne}{HTML}{419041}
  % \definecolor{colorTwo}{HTML}{cccccc}
  \definecolor{colorTwo}{HTML}{dddddd}
  \definecolor{colorThree}{HTML}{F1B52D}
  % \definecolor{colorThree}{HTML}{EFA126}
}

\usetheme{Default}
\usetitlestyle{Empty}
\usecolorstyle[colorPalette=grassColorPalette]{Britain}
\colorlet{backgroundcolor}{white}

\title{
\Huge
\textcolor{titleTextColor}{
\textsf{
% \textbf{
\fontsize{130}{100}\selectfont
\textbf{GRASS}\,{\firalight GIS}: A General-purpose Geospatial Research Tool
% }
}
}
}

@@

@@

\author{
Helena Mitasova\inst{1},
Vaclav Petras\inst{1}\,*\hspace{-7pt},
Anna Petrasova\inst{1},
\&
Markus Neteler\inst{2}
\\
% AGU allows to specify a scientific team, but it does not seem to fit
% with our case 100%; simply using formatting which is in the AGU program
% (but adding note)
Scientific Team: GRASS GIS Development Team**
}
\institute{
\large
\instlist{1}Center for Geospatial Analytics, North Carolina State University, USA;
\instlist{2}mundialis GmbH \& Co. KG, Germany;
*Corresponding author: wenzeslaus@gmail.com, vpetras@ncsu.edu;
% using Black Duck Open Hub to ``guestimate'' that
**Over 10 other members of the core team and numerous other contributors
\\[1.7cm]
\includegraphics[height=3cm]{ncstate}%
\hspace{1cm}%
\includegraphics[height=3cm]{cga}%
\hspace{3cm}%
\includegraphics[height=3.4cm]{mundialis}%
}

\hypersetup
{
    pdfauthor={H. Mitasova, V. Petras, A. Petrasova, M. Neteler},
    pdfsubject={AGU Fall Meeting 2018 Poster},
    pdftitle={GRASS GIS: A General-purpose Geospatial Research Tool},
    pdfkeywords={GIS, algorithms, methods, preservation, science, reproducibility}
}

@@

\newcommand{\blocktitlewrap}[1]{\textnormal{\textsf{\textsc{\huge#1}}}}
% it is not possible (?) to change block title in the class, using wrapper
% the command introduced using:
%   sed -i 's/\\block{\([^}]*\)}/\\block{\\blocktitlewrap{\1}}/g' main.tex

% bullet point style
\renewcommand{\labelitemi}{\textcolor{gray}{$\bullet$}\hspace{0.5ex}}

@@

\begin{document}

\node[above left,opacity=0.99,inner sep=0pt,outer sep=4cm] at (bottomleft -| topright)%
  {\includegraphics[width=0.3\paperwidth]{grass}};

\maketitle[width=0.92\textwidth]

% \maketitle
% \addlogo[north west]{(2,-1)}{9cm}{images/Grass_GIS}
%Please insert your institution logo here
% \addlogo[north east]{(-2,-2.5)}{4cm}{images/logo_FEM_CRI}
% \addlogo[north east]{(-2,-5.5)}{4cm}{images/NC_State_Seal}
% \addlogo[north east]{(-8,-2.5)}{4cm}{images/Logo_cvut}
% \addlogo[north east]{(-8,-6.5)}{4cm}{images/IWMI_logo}
% \addlogo[north east]{(-2,-10.5)}{4cm}{images/logo_ec-jrc}

@@

% Abstract
%
% GRASS GIS (grass.osgeo.org) is an open source software for geospatial analysis,
% remote sensing, general geoprocessing and visualization. It is a
% community-driven project with over 35 years of continuous software development
% based on scientific expertise from many geospatial fields. It is characterized
% by long-term releases, stable APIs, and emphasis on science. GRASS GIS
% distribution strives to provide single integrated environment for 2D and 3D
% raster analysis, image processing, vector data analysis, and spatio-temporal
% data processing. It supports large raster files (billions of cells), vector
% topology, coupling with databases, and 64 bit memory. New code based on recent
% research is typically contributed to GRASS GIS Addons repository. Mature, widely
% used code is then moved to the main code base to maximize integration and
% availability. Whether it is addons or the main code base, code is usually
% maintained by the community and preserved in long term even in cases when the
% original author no longer supports the code. To support the needs of scientists,
% the documentation includes not only links to the source code, its history, and
% its authors, but also links to research papers that describe the algorithms
% implemented in the modules.
%
% Code is not only maintained but also extended and improved. For example,
% watershed and stream extraction using least cost path approach was implemented
% in 1989 and extended for massive datasets in 2011. Similarly, vector topology
% cleaning was introduced in 2002, updated over time and substantially improved in
% 2016. Another example is a solar energy module available since 1993 and
% parallelized in 2017. The current stable version of GRASS GIS 7.4 provides new
% features for geosciences such as temporal framework with temporal algebra for
% large time-series processing, fusion of elevation models from various sensors,
% 3D flows, landform detection, image segmentation, simplified batch processing,
% and integrated Python editor among others. GRASS GIS runs in various
% environments including Linux, Mac, Windows, Docker, Raspberry Pi, and on HPC
% clusters. The modules are written in Python, C, or C++. Besides command line
% interface, GRASS GIS provides a graphical user interface, a Python API, and a C
% API. Interfaces with other languages such as R and Ruby are supplied by
% collaborating communities.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Overview}}{

\Large

\begin{itemize}
% Basic Topics
 \item Community-driven project
 \item 35 years of continuous software development
 \item long-term releases, stable APIs, and emphasis on science
 \item single integrated environment for 2D and 3D raster analysis, image processing, vector data analysis, and spatio-temporal data processing
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Functionality}}{

\Large

GRASS GIS
distribution strives to provide single integrated environment for 2D and 3D
raster analysis, image processing, vector data analysis, and spatio-temporal
data processing. It supports large raster files (billions of cells), vector
topology, coupling with databases, and 64 bit memory.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Code Life Cycle}}{

\Large

New code based on recent
research is typically contributed to GRASS GIS Addons repository. Mature, widely
used code is then moved to the main code base to maximize integration and
availability. Whether it is addons or the main code base, code is usually
maintained by the community and preserved in long term even in cases when the
original author no longer supports the code. To support the needs of scientists,
the documentation includes not only links to the source code, its history, and
its authors, but also links to research papers that describe the algorithms
implemented in the modules.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Spatio-Temporal Data Analysis}}{
The time dimension was introduced in GRASS GIS 7.0 for vector maps, rasters, and 3D rasters
which transformed GRASS GIS into a fully-featured temporal GIS \citep{Gebbert20141, gebbert2015grass}.
Time series of map layers are managed in space time datasets, a new data type in GRASS GIS,
and are still accessible also as individual map layers.
Based on the GRASS GIS Temporal Framework Python application programming interface (API),
more than 50 modules were implemented to manage, analyze, process
and visualize space time datasets.
More than 100,000 map layers can be now handled efficiently in GRASS GIS.
%
Example usages of this functionality include
analysis of the
European Climate Assessment \& Dataset ECA\&D \citep{Haylock2008_climate_series}
and temperate climate zone in the European Union identification \citep{Gebbert20141}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Water, Floods and Erosion}}{

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Vector Network Analysis}}{
In 2003 Radim Blazek introduced shortest path analysis (\gmodule{v.net.path}),
traveling salesman (\gmodule{v.net.salesman}),
and several other modules for vector network analysis (staring with \gmodule{v.net}).
Over the years, the number of available algorithms increased to over 15 modules
including, for example, centrality measures and connected components added in 2009 by Daniel Bundala.
In 2014 Stepan Turek implemented turns support into all relevant vector network modules.
Also since GRASS GIS 7.0 all vector network analysis tools provide fine control over node costs.
% combination with r.cost/r.walk workflows \citep{Petrasova2014}

\centering
\begin{minipage}{0.9\linewidth}
\begin{center}
\includegraphics[width=\textwidth]{network_trips}
A series of trips to collect samples across North Carolina, USA
for water quality measurements
as part of the Polyfluorinated Alkyl Substance Testing (PFAST) project.
Trips are distinguished by color and were computed using multiple
iterations of traveling salesman (\gmodule{v.net.salesman}).
\end{center}
\end{minipage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Image Segmentations}}{
\gamodule{r.smooth.seg} (formerly \gmodulenolink{r.seg})
was created by Alfonso Vitti \citep{vitti2008free, vitti2012mumford}
and implemented a piece-wise smooth approximation of the original image
according to \cite{mumford1989optimal} and \cite{march1997variational}
which can be used to reduce noise in the original image.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Lidar Data Processing}}{

@@

\node[above left,opacity=0.5,inner sep=0pt,outer sep=4cm] at (bottomleft -| topright)%
  {\includegraphics[width=0.3\paperwidth]{grass}};
